Grove: A Convergent Collaborative Structure-Editor Calculus

### 2021-05-11 grove-paper-3-formalism.tex
move term syntax and arity figures into new files and deduplicate
rename expr, patt, type to erecomp, precomp, trecomp
add decomp back to main formalism section (deduped)
make decomp return a pair and recomp accept a pair
finish decomp / recomp edge tracking

\begin{figure}
  \[
    \arraycolsep=0pt
    \begin{array}{lrlll}
      q    & {}\in Pat & {}::={} & x^{\id{u}} \mid \hole \mid \conflictHole{q,\ldots,q} \mid \multiVertex{u} \mid \cycleVertex{u} \\
      \tau & {}\in Typ & {}::={} & \tau \rightarrow^{\id{u}} \tau \mid Num^{\id{u}} \mid \hole \mid \conflictHole{\tau,\ldots,\tau} \mid \multiVertex{u} \mid \cycleVertex{u} \\
      e    & {}\in Exp & {}::={} & x^{\id{u}} \mid \lambda^{\id{u}} q \mathord{:}\tau.e \mid (e~e)^{\id{u}} \mid n^{\id{u}} \\
           &           &         & ~\mid e~\texttt{+}^{\id{u}}~e \mid e~\texttt{*}^{\id{u}}~e \mid \hole \mid \conflictHole{e,\ldots,e} \mid \multiVertex{u} \mid \cycleVertex{u} \\
    \end{array}
  \]
  \caption{Base syntax, as a grammar, for the lambda calculus that we are considering.}
  \label{fig:Base syntax}
\end{figure}

@@

\begin{figure}
  \[
    \arraycolsep=0pt
    \begin{array}{ll}
      \arity : \K \hfil \rightarrow & \wp(\P \times \{Exp, Pat, Typ\}) \\
      \hline
      \arity(\Root)={} & \left\{ (\Root, Exp) \right\} \\
      \arity(\PatVar(x))={} & \left\{ \right\} \\
      \arity(\ExpVar(x))={} & \left\{ \right\} \\
      \arity(\ExpLam)={} & \left\{ (\LamParam, Pat), (\mathtt{Type}, Typ), (\mathtt{Body}, Exp) \right\} \\
      \arity(\ExpApp)={} & \left\{ (\mathtt{Fun}, Exp), (\mathtt{Arg}, Exp) \right\} \\
      \arity(\ExpPlus)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
      \arity(\ExpTimes)={} & \left\{ (\mathtt{L}, Exp), (\mathtt{R}, Exp) \right\} \\
      \arity(\ExpNum(n))={} & \left\{ \right\} \\
      \arity(\TypArrow)={} & \left\{ (\mathtt{Arg}, Typ), (\mathtt{Result}, Typ) \right\} \\
      \arity(\TypNum)={} & \left\{ \right\} \\
    \end{array}
  \]
  \caption{Constructors, Indexes and Arity}
  \label{fig:Constructors, Indexes and Arity}
\end{figure}

@@

A grove is a 3-tuple of the following form:
\[
  (NP, MP, U, D)
\]
where
\begin{align*}
  % e_r &= \expr(v_0) \\
  NP &= \{\expr(v) : |\parents(v)| = 0\} \\
  MP &= \{\expr(v) : |\parents(v)| > 1\} \\
  U  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\} \\
  D  &= \{\e : G(\e) = \Minus\}
\end{align*}
For example, the grove corresponding to the graph in \autoref{fig:Decomposition example graph} is
given in \autoref{fig:Decomposition example grove}.

@@

The third component $U$ is the set of expressions derived from the remaining vertices which necessarily have exactly one parent.
These vertices form structures formally called unicycular graphs~\citep{DBLP:journals/algorithmica/KruskalRS90}, or \emph{unicycles} for short.
Every unicycle consists of a single cycle where each vertex in the cycle may have additional children that are not part of the cycle.
Breaking the cycle turns a unicyclic graph into a tree.
% TODO: try clearing up "the root" as e.g., "the component root" or "the root of this components of the grove"
% TODO: make sure we say "the graph root" when talking about the actual root, and "the component root" everywhere else.
To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id, $u$, as the root of that tree.
Formally, we choose the vertex $v$ such that $v = \min(\ancestors(v))$ because
only a vertex on the cycle can be its own ancestor.
When an edge to this vertex appears, we place a reference to it $\cycleVertex{u}$.
In our example, ${+}_{\id{46}}$ is referenced by the $LEFT$ edge of ${*}_{\id{48}}$.

>>>>

\figureTermSyntax

@@

\figureArity

@@

A grove is a 3-tuple of the following form:
\[
  (NP, MP, U)
\]
where
\begin{align*}
  \figureDecompositionDefDecompComponents
\end{align*}
For example, the grove corresponding to the graph in \autoref{fig:Decomposition example graph} is
given in \autoref{fig:Decomposition example grove}.

\figureDecompositionDef

\figureDecompositionDefHelpers

@@

The third component $U$ is the set of expressions derived from the remaining vertices which necessarily have exactly one parent.
These vertices form structures formally called unicycular graphs~\citep{DBLP:journals/algorithmica/KruskalRS90}, or \emph{unicycles} for short.
Every unicycle consists of a single cycle where each vertex in the cycle may have additional children that are not part of the cycle.
Breaking the cycle turns a unicyclic graph into a tree.
% TODO: try clearing up "the root" as e.g., "the component root" or "the root of this components of the grove"
% TODO: make sure we say "the graph root" when talking about the actual root, and "the component root" everywhere else.
To break the cycle, we arbitrarily choose the vertex on the cycle with the smallest id, $u$, as the root of that tree.
Formally, we choose the vertex $v$ such that $v = \min{\ancestors{v}}$ because
only a vertex on the cycle can be its own ancestor.
When an edge to this vertex appears, we place a reference to it $\cycleVertex{u}$.
In our example, ${+}_{\id{46}}$ is referenced by the $LEFT$ edge of ${*}_{\id{48}}$.
